A Global Picture of Food Insecurity and Hunger

### 2020-10-22 docs/modeling_fies.tex
Updates from Overleaf

@@

We observed for the moderate-to-severe as well as the severe RF regression model an in-sample mean squared error (MSE) between 0.002-0.005, with a R-squared (${R}^2$) value of 0.999 (See Fig. \ref{fig:rf_in-sample}). To test the out-of-sample prediction performance we used ten-fold cross-validation. First, we partitioned the data into ten equal sized subsamples. Once partitioned, we trained the RF regression models on nine subsamples and tested the models on the remaining one. This process gets repeated ten times, with each subsample being used once for testing. We then calculated the average MSE and ${R}^2$ over all ten out-of-sample predictions. Similarly to the in-sample model validation metrics, the average out-of-sample MSE is between 0.008-0.016, with an average ${R}^2$ of around 0.99 for both models (See Fig. \ref{fig:rf_out-sample}). Besides using the full set of countries to validate the out-of-sample performance, we halved the sample size by excluding high-income countries (HIC) (according to the World Bank definition, REFERENCE NEEDED) and performed ten-fold cross-validation for this limited set of countries. We found that even when excluding HIC the average out-of-sample MSE is between 0.013-0.023, with an average ${R}^2$ of around 0.98 for both models.

To get the best possible model performance, it is necessary to tune the hyperparamters, which control the process in which the random forest algorithm is run.  Key hyperparameters include the number of decision trees to generate, the number of variables randomly selected as candidates for splitting a node, and the average number of observations in a leaf node.  We calculated the Out-Of-Bag (OOB) error for different combinations of hyperparameters for the moderate-to-severe and severe models. By choosing the combination of hyperparameters with the smallest OOB error we found the optimal combination for both models respectively. We found that 5000 trees are more than sufficient for the error rates to converge (See Fig. \ref{fig:rf_error}).

>>>>

%do we need to add a section where we state which fies data we use? like countries and years?

@@

We observed for the moderate-to-severe as well as the severe RF regression model an in-sample mean squared error (MSE) between 0.002-0.005, with a R-squared (${R}^2$) value of 0.999 (See Fig. \ref{fig:rf_in-sample}). To test the out-of-sample prediction performance we performed ten-fold cross-validation. First, we partitioned the data into ten equal sized subsamples. Once partitioned, we trained the RF regression models on nine subsamples and tested the models on the remaining one. This process gets repeated ten times, with each subsample being used once for testing. We then calculated the average MSE and ${R}^2$ over all ten out-of-sample predictions. Similarly to the in-sample model validation metrics, the average out-of-sample MSE is between 0.008-0.016, with an average ${R}^2$ of around 0.99 for both models (See Fig. \ref{fig:rf_out-sample}). Besides using the full set of countries to validate the out-of-sample performance, we halved the sample size by excluding high-income countries (HIC) (according to the World Bank definition, REFERENCE NEEDED?) and performed ten-fold cross-validation for this limited set of countries. We found that excluding HIC does not change the out-of-sample performance of the models. The average out-of-sample MSE is between 0.013-0.023 with an average ${R}^2$ of around 0.98 for both models.

To get the best possible model performance, it is necessary to tune the hyperparamters, which control the process in which the random forest algorithm is run.  Key hyperparameters include the number of decision trees to generate, the number of variables randomly selected as candidates for splitting a node, and the average number of observations in a leaf node. For the latter two we calculated the Out-Of-Bag (OOB) error for different combinations of hyperparameters for the moderate-to-severe and severe models. By choosing the combination of hyperparameters with the smallest OOB error we found the optimal combination for both models respectively. For the number of decision trees we found that 5000 trees are more than sufficient for the error rates to converge (See Fig. \ref{fig:rf_error}).
